The CrypTool Script: \\ Cryptography, Mathematics and More \\ \vspace{30pt} {\large Background reading\\%*[-8pt] for the free e-learning program CrypTool\\ (10th edition -- distributed with CrypTool version 1.4.30)} \vspace{60pt}

### 2013-01-14 script/de/chapters/primes.tex
Several minor changes (for consistency and correctness), mostly in chap 4.

\item Das Benfordsche Gesetz\label{primes:Benford-Law}\footnote{%
        \url{http://de.wikipedia.org/wiki/Benfordsches_Gesetz}, \\
        \url{http://www.spiegel.de/wissenschaft/mensch/0,1518,632541,00.html},\\
        \url{http://arxiv.org/PS_cache/arxiv/pdf/0906/0906.2789v1.pdf}. 
        }$^,$\footnote{%
        Didaktische Darstellungen zu Anwendungen von Benfords Gesetz finden sich
        unter:
        % \begin{itemize}
        \begin{compactitem}
         \item Rüdeger Baumann: \glqq Ziffernanalyse zwecks Betrugsaufdeckung
              --- Beispiel für kompetenzorientierten und kontextbezogenen
              Informatikunterricht\grqq, \\
              in LOGIN, Informatische Bildung und Computer in der Schule,
              Nr. 154/155, 2008, S. 68-72
         \item Norbert Hungerbühler: \glqq Benfords Gesetz über führende
              Ziffern\grqq, \\
              März 2007,
              \url{http://www.educ.ethz.ch/unt/um/mathe/ana/benford}
        \end{compactitem}
        % \end{itemize}
        \vspace{-\baselineskip} % TODO: Geht das nicht anders?
        }
     gilt nicht für Primzahlen.

@@

\subsubsection*{Primzahlsatz}
Die Anzahl der Primzahlen PI(x) bis zu einer gegebenen Zahl $x$ kann durch
eine Formel nach dem sogenannten Primzahlsatz näherungsweise bestimmt werden
(siehe Kapitel chapter \ref{l_Primes_Distrib-of-Primes}).
$PI(x)$ bezeichnet die Anzahl der Primzahlen kleiner oder gleich der Zahl $x$.
Dann lautet die Formel
\[PI(x)\sim \frac{x}{\ln{x}}.\]
\index{PI(x), $\Pi(x)$}
Es ist zu beachten, dass diese Formel die Anzahl der Primzahlen kleiner oder gleich $x$ nur
ungefähr angibt. Sie wird jedoch genauer je größer $x$ wird.
Im folgenden wollen wir den Primzahl"-satz nutzen, um uns die Verteilung der Primzahlen anzusehen.

>>>>

\item Das Benfordsche Gesetz\label{primes:Benford-Law}\footnote{%
        \url{http://de.wikipedia.org/wiki/Benfordsches_Gesetz}, \\
        \url{http://www.spiegel.de/wissenschaft/mensch/0,1518,632541,00.html},\\
        \url{http://arxiv.org/PS_cache/arxiv/pdf/0906/0906.2789v1.pdf}. 
        }$^,$\footnote{%
        Didaktische Darstellungen zu Anwendungen von Benfords Gesetz finden sich
        unter:
        % \begin{itemize}
        \begin{compactitem}
         \item Rüdeger Baumann: \glqq Ziffernanalyse zwecks Betrugsaufdeckung
              --- Beispiel für kompetenzorientierten und kontextbezogenen
              Informatikunterricht\grqq, \\
              in LOGIN, Informatische Bildung und Computer in der Schule,
              Nr. 154/155, 2008, S. 68-72
         \item Norbert Hungerbühler: \glqq Benfords Gesetz über führende
              Ziffern\grqq, \\
              März 2007,
              \url{http://www.educ.ethz.ch/unt/um/mathe/ana/benford}
        \end{compactitem}
        % \end{itemize}
        %\vspace{-\baselineskip} % TODO: Geht das nicht anders? Nützte nichts.
        }%\vspace{-\baselineskip} Nützte nichts.
     gilt nicht für Primzahlen.

@@

\subsubsection*{Primzahlsatz}
Die Anzahl der Primzahlen PI(x) bis zu einer gegebenen Zahl $x$ kann durch
eine Formel nach dem sogenannten Primzahlsatz näherungsweise bestimmt werden
(siehe Kapitel \ref{l_Primes_Distrib-of-Primes}).
$PI(x)$ bezeichnet die Anzahl der Primzahlen kleiner oder gleich der Zahl $x$.
Dann lautet die Formel
\[PI(x)\sim \frac{x}{\ln{x}}.\]
\index{PI(x), $\Pi(x)$}
Es ist zu beachten, dass diese Formel die Anzahl der Primzahlen kleiner oder gleich $x$ nur
ungefähr angibt. Sie wird jedoch genauer je größer $x$ wird.
Im folgenden wollen wir den Primzahl"-satz nutzen, um uns die Verteilung der Primzahlen anzusehen.

### 2013-01-14 script/en/chapters/cryptomethods.tex
Several minor changes (for consistency and correctness), mostly in chap 4.

Based on those assumptions, the effort to find a AES-256 key is reduced to
$2^{119}$ time and $2^{77}$ memory. In the case of AES-192 the attack is even less
practical, for AES-128 the authors do not provide an attack.

@@

% --------------------------------------------------------------------------
\section[Asymmetric encryption]
{Asymmetric encryption\footnotemark}
  \footnotetext{%
    With CT1\index{CrypTool 1} you can execute RSA encryption and decryption
    (using the menu path {\bf Crypt \textbackslash{} Asymmetric}).
    In both cases you must select a RSA key pair. Only in the case of decryption
    the secret RSA key is necessary -- so here you are asked to enter the PIN.\\
    With CT2\index{CrypTool 2} you can also perform asymmetic methods
    (using the templates under {\bf Cryptography \textbackslash{} Modern}).\\
    JCT\index{JCrypTool} offers asymmetic methods like RSA both within
    the {\bf Visuals} menu of the Default Perspective as well as within the
    Algorithm Perspective.
  }
\nopagebreak
In the case of {\em asymmetric} encryption \index{Encryption!asymmetric} each
subscriber has a personal pair of keys consisting of a {\em secret}
\index{Key!secret} key and a {\em public} key\index{Key!public}. The public
key, as its name implies, is made public, e.g. in a key directory on the
Internet (this kind of ``bill-board'' is also called just directory or
public key ring).\par \vskip + 3pt

>>>>

Based on those assumptions, the effort to find a AES-256 key is reduced to
$2^{119}$ time and $2^{77}$ memory (considering asymmetric complexity).
In the case of AES-192 the attack is even less practical, for AES-128
the authors do not provide an attack.

@@

% --------------------------------------------------------------------------
\section[Asymmetric encryption]
{Asymmetric encryption\footnotemark}
  \footnotetext{%
    With CT1\index{CrypTool 1} you can execute RSA encryption and decryption
    (using the menu path {\bf Crypt \textbackslash{} Asymmetric}).
    In both cases you must select a RSA key pair. Only in the case of decryption
    the secret RSA key is necessary -- so here you are asked to enter the PIN.\\
    With CT2\index{CrypTool 2} you can also perform asymmetric methods
    (using the templates under {\bf Cryptography \textbackslash{} Modern}).\\
    JCT\index{JCrypTool} offers asymmetric methods like RSA both within
    the {\bf Visuals} menu of the Default Perspective as well as within the
    Algorithm Perspective.
  }
\nopagebreak
In the case of {\em asymmetric} encryption \index{Encryption!asymmetric} each
subscriber has a personal pair of keys consisting of a {\em secret}
\index{Key!secret} key and a {\em public} key\index{Key!public}. The public
key, as its name implies, is made public, e.g. in a key directory on the
Internet (this kind of ``bill-board'' is also called just directory or
public key ring).\par \vskip + 3pt
